---
title: "Lab 6"
author: "Thorax Hembleecker"
date: "20XX-03-19"
output: pdf_document
---

## 2. Confidence intervals

``` {r 2a}
data(infert)
n = length(infert$age) # Sample size.
sigma = 5
```

``` {r 2b}
mean(infert$age) + qnorm(0.975)*sigma/sqrt(n) # Upper bound
mean(infert$age) - qnorm(0.975)*sigma/sqrt(n) # Lower bound
```

``` {r 2c}
mean(infert$age) + qnorm(0.995)*sigma/sqrt(n) # Upper bound
mean(infert$age) - qnorm(0.995)*sigma/sqrt(n) # Lower bound
```

``` {r 2d}
mean(infert$age) + qnorm(0.95)*sigma/sqrt(n) # Upper bound
mean(infert$age) - qnorm(0.95)*sigma/sqrt(n) # Lower bound
```

``` {r 2e}
mean(infert$age) + qnorm(0.95)*sigma/sqrt(n) # Upper bound
```

``` {r 2f}
mean(infert$age) - qnorm(0.99)*sigma/sqrt(n) # Lower bound
```

## 3. t tests

``` {r 3a}
mean(infert$age) + qt(0.975, df=n-1)*(sd(infert$age)/sqrt(n)) # Upper bound
mean(infert$age) - qt(0.975, df=n-1)*(sd(infert$age)/sqrt(n)) # Lower bound

t.test(infert$age)
```

``` {r 3b}
mean(infert$age) + qt(0.995, df=n-1)*(sd(infert$age)/sqrt(n)) # Upper bound
mean(infert$age) - qt(0.995, df=n-1)*(sd(infert$age)/sqrt(n)) # Lower bound

t.test(infert$age, conf.level=0.99)
```

``` {r 3c}
mean(infert$age) + qt(0.95, df=n-1)*(sd(infert$age)/sqrt(n)) # Upper bound

t.test(infert$age, alternative="less")
```

``` {r 3d}
mean(infert$age) - qt(0.95, df=n-1)*(sd(infert$age)/sqrt(n)) # Lower bound

t.test(infert$age, alternative="greater")
```

``` {r 3e}
mean(infert$age) + qt(0.99, df=n-1)*(sd(infert$age)/sqrt(n)) # Upper bound

t.test(infert$age, conf.level=0.99, alternative="less")
```

``` {r 3f}
mean(infert$age) - qt(0.99, df=n-1)*(sd(infert$age)/sqrt(n)) # Lower bound

t.test(infert$age, conf.level=0.99, alternative="greater")
```

# Hypothesis Testing

## 4.

``` {r 4a}
z.star = (15.8-16)/(0.4/sqrt(40)) # Test statistic.
pnorm(z.star)
```

``` {r 4b}
z.star = (16.2-16)/(0.4/sqrt(40)) # Test statistic.
1 - pnorm(z.star)
```

``` {r 4c}
z.star = (191-166.3)/(50/sqrt(30)) # Test statistic.
2*pnorm(-abs(z.star))
```

## 5.

``` {r 5a}
t.star = (19.9-25.2)/(19.8/sqrt(52)) # Test statistic.
pt(t.star, df=51) # df=n-1
```

``` {r 5b}
t.star = (32.9-25.2)/(19.8/sqrt(52)) # Test statistic.
1 - pt(t.star, df=51)
```

``` {r 5c}
t.star = (82.3-71.2)/(34.2/sqrt(40)) # Test statistic.
2*pt(-abs(t.star), df=39)
```

## 6.

```{r 6a}
avocado = read.csv("avocado.csv")
```

$\ H_0: \mu = 1.7$ vs. $\ H_A: \mu \neq 1.7$
```{r 6b}
t.test(avocado$price, mu=1.7)
```
Interpretation: Since the p-value = 0.6046 is greater than the significance level, $\alpha$ = 0.05, we fail to reject the null hypothesis. We cannot conclude that the average price of avocado is different from \$1.70.
We estimate that the price of avocado is on average \$1.732.
I am 95% confident that the interval (1.6085, 1.8556) contains the true average price of avocado.

$\ H_0: \mu <= 1.3$ vs. $\ H_A: \mu > 1.3$
```{r 6c}
t.test(avocado$price, mu=1.3, alt="greater", conf.level=0.99)
```
Interpretation: Since the p-value = 2.326e-09 is less than the significance level, $\alpha$ = 0.01, we succeed in rejecting the null hypothesis. We conclude that the average price of avocado is greater than \$1.30.
We estimate that the price of avocado is on average \$1.732.
I am 99% confident that the interval (1.5843, infinity) contains the true average price of avocado.

$\ H_0: \mu >= 2$ vs. $\ H_A: \mu < 2$
```{r 6d}
t.test(avocado$price, mu=2, alt="less", conf.level=0.90)
```
Interpretation: Since the p-value = 3.177e-05 is less than the significance level, $\alpha$ = 0.1, we succeed in rejecting the null hypothesis. We conclude that the average price of avocado is less than \$2.00.
We estimate that the price of avocado is on average \$1.732.
I am 90% confident that the interval (-infinity, 1.812) contains the true average price of avocado.

# Power and sample size

## 7.

```{r 7a}
sd = 0.4
n = 36
alpha = 0.05
mu0 = 16 # Null hypothesis
mu1 = 15.8 # Alternative hypothesis
```

```{r 7b}
x = mu0 + qnorm(alpha)*sd/sqrt(n)
pnorm((x-mu1)/(sd/sqrt(n)))
```

```{r 7c}
mu1 = seq(15.65, 16, 0.001)
```

```{r 7d}
n = 36
x = mu0 + qnorm(alpha)*sd/sqrt(n)
plot(mu1, pnorm((x-mu1)/(sd/sqrt(n))), type="l", ylab="Power", xlab=expression(mu[1]) )

n = 50
x = mu0 + qnorm(alpha)*sd/sqrt(n)
lines(mu1, pnorm((x-mu1)/(sd/sqrt(n))), col=2)
```

```{r 7e}
alpha = 0.05
beta = 0.05
mu0 = 16
mu1 = 15.8
sd = 0.4

ceiling((((qnorm(alpha/2)+qnorm(beta))*sd)/(mu1-mu0))^2) # 2-tailed
```

```{r 7f}
alpha = 0.01
beta = 0.1
mu0 = 16
mu1 = 15.8
sd = 0.4

ceiling((((qnorm(alpha)+qnorm(beta))*sd)/(mu1-mu0))^2)
```
