---
title: "Lab 10"
author: "Thorax Hembleecker"
date: "20XX-04-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.

$H_0: p_1 = 0.2, p_2 = 0.3, p_3 = 0.15, p_4 = 0.1, p_5 = 0.25$ vs. $H_1:$ at least one of these equalities does not hold
```{r 2a-c}
n = 300
obs1 = 62 # Brown hair
obs2 = 92 # Black hair
obs3 = 50 # Blonde hair
obs4 = 25 # Orange hair
obs5 = 71 # Other hair

p1 = 0.2 # Hypothesized proportion for brown hair
exp1 = n*p1 # Expected count for brown hair
cont1 = (obs1 - exp1)^2 / exp1 # Contribution to test statistic
p2 = 0.3
exp2 = n*p2
cont2 = (obs2 - exp2)^2 / exp2
p3 = 0.15
exp3 = n*p3
cont3 = (obs3 - exp3)^2 / exp3
p4 = 0.1
exp4 = n*p4
cont4 = (obs4 - exp4)^2 / exp4
p5 = 0.25
exp5 = n*p5
cont5 = (obs5 - exp5)^2 / exp5

x2.star = cont1 + cont2 + cont3 + cont4 + cont5
x2.star
```

```{r 2d}
pval = 1 - pchisq(x2.star, df=4)
pval
pval < 0.05
```
The p-value, 0.788, is greater than the significance level, 0.05, so we fail to reject the null hypothesis and conclude that our estimated proportions are not significantly different from the actual proportions.

## 3.

```{r 3}
categories = c(62, 92, 50, 25, 71)
nullprob = c(0.2, 0.3, 0.15, 0.1, 0.25)
chisq.test(categories, p=nullprob) # Goodness-of-fit test
```
The test statistic and p-value produced by chisq.test match the ones produced manually.

## 4.

```{r 4}
library(vcd)
data(Arthritis)

categories = table(Arthritis$Improved)
nullprob = c(1/3, 1/3, 1/3)
chisq.test(categories, p=nullprob)
```
The p-value, 0.000912, is less than the significance level, 0.01, so we reject the null hypothesis and conclude that at least one of the true proportions differs from 1/3 (really at least two, since the proportions must add up to 1). Therefore, either marked improvement, no improvement, or some improvement must be more likely than the other options.

## 5.

$H_0: Improved$ is independent of $Treatment$ vs. $H_1: Improved$ is not independent of $Treatment$
```{r 5}
chisq.test(table(Arthritis$Improved, Arthritis$Treatment), correct=F)
```
The p-value, 0.001, is less than the significance level, 0.05, so we can reject the null hypothesis and conclude that $Improved$ is not independent of $Treatment$, meaning that arthritis treatments do have an impact on whether the patient's symptoms improve.

## 6.

$H_0:$ Having a college degree is not associated with political leanings vs. $H_1:$ Having a college degree is associated with political leanings
```{r 6a-b}
datamat = matrix(c(90, 120, 40, 140, 65, 45), ncol=3, byrow=TRUE)
datamat

colnames(datamat) = c("Rep", "Dem", "Ind")
rownames(datamat) = c("Yes", "No")
datatab = as.table(datamat)
datatab
```

```{r 6c}
chisq.test(datatab, correct=F)
```
The p-value, 1.06e-6, is less than the presumed significance level of 0.05, so we can reject the null hypothesis and conclude that having a college degree is associated with political leanings.