---
title: "Lab 1"
author: "Thorax Hembleecker"
date: "20XX-01-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r importData, include=FALSE}
arthritis = read.csv("arthritis.csv")
```

## Case C Numerical Summaries and Graphics

### Numerical summaries

``` {r caseC1_Num}
tab1 = table(arthritis$Improved) # tab1 -> absolute frequency
tab1
prop.table(tab1) # Relative frequency ("prop" as in proportional)
```

### Graphical summaries

``` {r caseC1_Graph}
barplot(tab1) # Absolute frequency bar graph
barplot(prop.table(tab1)) # Relative frequency bar graph
```

Observations:

1. The "None" type is the most common in Improvement :(

2. "Some" is the least common

3. It's weird that people either have no improvement or marked improvement

### Summarizing treatment variable

``` {r caseC2}
tab2 = table(arthritis$Treatment)
tab2
prop.table(tab2)

barplot(tab2)
barplot(prop.table(tab2))
```

## Case Q Numerical Summaries and Graphics

### Numerical summaries

``` {r caseQ1}
## Case Q summary for Age variable:

# Centers:
mean(arthritis$Age)
median(arthritis$Age)
library(modeest) # Library for mode function
mlv(arthritis$Age, method = "mfv") # Mode

# Variablility:
sd(arthritis$Age) # Standard deviation
var(arthritis$Age) # Variance
sqrt(var(arthritis$Age)) # Calculates S.D. from variance

quantile(arthritis$Age, 0.25) # Q1
quantile(arthritis$Age, c(0.25, 0.75)) # Q1 and Q3
IQR(arthritis$Age) #IQR

fivenum(arthritis$Age) # Five Number Summary
summary(arthritis$Age) # Similar, but includes mean
```

### Graphical summaries

``` {r caseQ1_Graph}
par(mfrow = c(1,2)) ## Creates a 1x2 matrix, apparently
# Frequency:
hist(arthritis$Age, main = "Histogram of Age", xlab = "Age", ylab = "Frequency", col = "aquamarine4")
# Density:
hist(arthritis$Age, main = "Histogram of Age", xlab = "Age", ylab = "Density", col = "aquamarine4", freq = FALSE)
par(mfrow = c(1, 1))

# x/ylab changes x/y-axis label; main changes title of plot

# Mean:
mean_age = mean(arthritis$Age)
mean_age
# Draw histogram...
hist(arthritis$Age, main = "Histogram of Age", xlab = "Age", col = "lightblue")
# ...then add line for mean
abline(v = mean_age, col = "red") # "v" is vertical intercept

boxplot(arthritis$Age, range = 0, main = "Standard (Skeletal) Boxplot of Age", ylab = "Age") # Standard/skeletal boxplot
boxplot(arthritis$Age, main = "Modified Boxplot of Age", ylab = "Age") # Modified boxplot

library(moments)
skewness(arthritis$Age)

qqnorm(arthritis$Age)
qqline(arthritis$Age)
```

The samples mostly follow the line, but trail off a little at each end.