---
title: "Lab 11"
author: "Thorax Hembleecker"
date: "20XX-04-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.

```{r 2a-b}
data(cars)
summary(cars)

plot(x=cars$speed, y=cars$dist, xlab="Speed", ylab="Stopping Distance")

# Correlation
r = cor(cars$speed, cars$dist)
r
```
There appears to be a positive linear relationship between speed and stopping distance. There are a couple points that seem like possible unusual points (e.g. (14, 80) and (24, 120)), but even those fit firmly within the overall trajectory.

```{r 2d-f}
x = cars$speed # Independent
y = cars$dist # Dependent

x.bar = mean(x)
y.bar = mean(y)
x.sd = sd(x)
y.sd = sd(y)

beta1 = r*(y.sd/x.sd)
beta1

beta0 = y.bar - beta1*x.bar
beta0
```
On average, the stopping distance will increase by approximately 3.93 feet for every increase of 1 mile per hour in speed. $\beta_0$ has no meaning in context because there will be no stopping distance when the car is not moving, and 0 is outside of our domain (where the minimum observed value is 4).

```{r 2g}
lm = lm(y~x)
summary(lm)
```
```{r 2h}
predict(lm, newdata=data.frame(x=21))
```
The prediction equation is y = 3.932409x - 17.57909, and the predicted value for x = 21 is 65.001.

```{r 2i}
plot(x, y, xlab="Speed", ylab="Stopping Distance")
abline(lm, col="red", lwd=2) # Line width 2.
```
## 3.

```{r 3a}
summary(lm)
cor(x, y)^2
```
$R^2$ is equal to 0.6511, as verified by squaring the correlation.

```{r 3b}
plot(lm, which=1) # Residual vs. fitted
plot(lm, which=2) # QQ plot
```
All three assumptions remain reasonable, although homoskedasticity may be questionable and the data appears somewhat right-skewed. The regression line is a good fit for the model, and 23, 35, and 49 are unusual points.

## 4.

```{r 4a}
confint(lm)
confint(lm, level=0.99)
```
```{r 4b}
sum.lm = summary(lm)
sum.lm

cor.test(y,x)
```
The tests show that stopping distance is indeed associated with speed.

## 5.

```{r 5}
predict(lm, newdata=data.frame(x=21), interval="confidence")
```
